### 2017-12-28 paper.Rmd
big paper update

---
title: "Silent pessimists: researchers' perceptions of malaria eradication"
bibliography: library.bib
output: 
  pdf_document:
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
# No scientific notation
# options(scipen=999)

# Packages (load child-specific packages in their respective files)
library(knitr)
library(xtable)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.align = 'center')
options(xtable.comment = FALSE)

# Get data
source('prepare_data.R', echo = FALSE)

theme_paper <- theme_publication
```

```{r}
authors <- data.frame(name = c('Joe Brew',
                               'Menno Pradhan',
                               'Jacqueline Broerse',
                               'Elisa Sicuri'),
                      email = c('joe@databrew.cc',
                                'm.p.pradhan@vu.nl',
                                'j.e.w.broerse@vu.nl',
                                'elisa.sicuri@isglobal.org'),
                      affilitation = c("isglobal,cism,vu",
                                       "vu,uva",
                                       "vu",
                                       "isglobal,icl,cism"),
                      footnote = c("Corresponding Author",
                                   rep('', 3)),
                      stringsAsFactors = FALSE)
institutions <- data.frame(code = c('isglobal',
                                    'icl',
                                    'cism',
                                    'vu',
                                    'uva'),
                           institution = c('Barcelona Institute for Global Health',
                                           'Imperial College London',
                                           'Centro de Investigação em Saúde de Manhiça',
                                           'VU University Amsterdam',
                                           'University of Amsterdam'),
                           address = c('c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain',
                                       'South Kensington Campus, London SW7 2AZ, U.K.',
                                       'Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique',
                                       'De Boelelaan 1105, 1081 HV Amsterdam, Netherlands',
                                       'REC E, Roetersstraat 11, Amsterdam, Netherlands'),
                      stringsAsFactors = FALSE)

make_author <- function(authors,
                        institutions){

  out <- c()
  for (i in 1:nrow(authors)){
    this_author <- authors$name[i]
    these_institutions <- authors$affilitation[i]
    these_institutions <- unlist(strsplit(these_institutions, ','))
    author_text <- paste0(this_author)
    for (j in 1:length(these_institutions)){
      this_institution <- these_institutions[j]
      if(any(grepl(this_institution, out))){
        author_text <- paste0(author_text,
                              '\\footrecall{',
                              this_institution,
                              '}'
                              )
      } else {
        # new institution, get full name
        full_name <- institutions$institution[institutions$code == this_institution]
        address <- institutions$address[institutions$code == this_institution]
        author_text <- paste0(author_text,
                              '\\footremember{',
                              this_institution,
                              '}{',
                              full_name,
                              ': ',
                              address,
                              '}')
      }
    }
    out[i] <- author_text
  }
  out <- paste0(out, collapse = ', ')
  cat(out)
}
```

\begin{center}
\begin{large}

```{r, results = 'asis', eval = TRUE}
make_author(authors = authors,
            institutions = institutions)
```

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Abstract}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

Quantifying an event's probability and time frame is essential for calculating its expected value. In the case of global malaria eradication, uncertainty regarding feasibility makes it difficult for policymakers and public health practitioners to make fully informed decisions. The opportunity cost of investments in eradication-specific interventions can be high, particularly in contexts with other urgent health priorities. In a systemic survey of malaria researchers, we query perceptions regarding the likelihood and time frame of eradication, as well as the perceived chief areas needed for improvement in order for eradication to be achieved. We assess pessimism/optimism (via the proxy of years-to-eradication), broken down by area of expertise, and adjust for selection bias. Our results show a disconnect between optimistic public institutional discourse and pessimistic private opinion, suggesting either (a) the necessity of a realistic accounting of eradication's low short-term likelihood in health planning and financing, or (b) the need for a more compelling case regarding eradication's feasibility to be made to the research community.
\end{changemargin}
\end{center}

# Introduction

## Background

Malaria is a parasitic disease transmitted to humans by mosquitoes. The _Plasmodium falciparum_ species, transmitted by the female _Anopheles_ mosquito, accounts for a large majority of the 200 million annual cases as well as the half million annual deaths [@White2014] [@World2016]. Malaria "elimination", the "interruption of all local transmission of the infection in a country or region" [@Tanner2015] is actively being pursued by dozens of countries around the world, leading to a renewed push for "eradication" (the global elmination of malaria) [@Rabinovich_2017].

This is not the first time that eradication has been in the international spotlight. The scientific and public health communities have had eradication on their policy agenda since the WHO established the Global Malaria Eradication Program in the 1950s [@Njera2011]. In 1957, U.S. President Dwight Eisenhower told Congress that malaria could be expected to be eradicated in five years. Following the failure of the WHO's first attempt, the focus shifted away from global eradication towards local elimination and control strategies.  In recent years, much of the discourse regarding malaria has shifted back to global eradication [@Roberts2007], with funders, researchers, and public health practitioners rallying to the cause [@Tanner2015]. The Bill and Melinda Gates Foundation has begun actively promoting eradication as feasible "within a generation" [@Gates2014], and the leadership of the World Health Organization (WHO) has set similarly ambitious goals, stating the objective of eliminating malaria in 65 new countries from 2015 through 2030 (Figure 1).

```{r, fig.cap='Countries with malaria: Observed and WHO objectives', fig.height = 3, fig.width = 4}
# Read WHO data (compiled manually from WHO website)
# Data from http://www.who.int/malaria/areas/elimination/overview/en/'
who <- readr::read_csv('who_elimination_data.csv')
who$year <- as.numeric(substr(who$year, 6, 9))
projection <- data.frame(year = seq(2030.2, 2035, length = 20),
                         countries = seq(33, 0, length = 20))
g <- ggplot(data = who %>% mutate(past = ifelse(past, 'Historical',
                                           'WHO objective')) %>%
         filter(!is.na(with_malaria)) ,
       aes(x = year,
           y = with_malaria,
           color = past)) +
  geom_line() +
  geom_point() +
  theme_paper() +
  labs(x = '',
       y = 'Countries') +
  scale_color_manual(name = '',
                     values = c('darkorange', 'blue')) +
  ylim(0, 130) +
  theme(axis.text.x = element_text(angle = 90)) #+
  # geom_line(data = projection,
  #           aes(x = year,
  #               y = countries,
  #               color = NA,
  #               group = NA),
  #           color = 'black',
  #           lty = 2) +
  # geom_hline(yintercept = 0, col = 'red') 
g
# #Make animation for presentation
# make_animation <- function(){
#   g <- 
#     ggplot(data = who %>% mutate(past = ifelse(past, 'Historical',
#                                            'WHO objective')) %>%
#          filter(!is.na(with_malaria)) ,
#        aes(x = year,
#            y = with_malaria,
#            color = past)) +
#   geom_line() +
#   geom_point() +
#   theme_black() +
#   labs(x = '',
#        y = 'Countries',
#        title = 'Countries with malaria: Observed and WHO objectives',
#        subtitle = 'Data from http://www.who.int/malaria/areas/elimination/overview/en/') +
#   scale_color_manual(name = '',
#                      values = c('darkorange', 'blue')) +
#   ylim(0, 130) +
#   theme(axis.text.x = element_text(angle = 90)) +
#   geom_hline(yintercept = 0, col = 'red') +
#     xlim(1970, 2060)
#   add_zero <- function (x, n) {
#     x <- as.character(x)
#     adders <- n - nchar(x)
#     adders <- ifelse(adders < 0, 0, adders)
#     for (i in 1:length(x)) {
#         if (!is.na(x[i])) {
#             x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""), 
#                 x[i], collapse = "")
#         }
#     }
#     return(x)
# }
#   counter <- 0
#   a <- (seq(1,5, length = 15))^2.5
#   a <- round(a)
#   a <- c(a, rev(a) +2)
#   for (i in a){
#     message(counter)
#     counter <- counter +1
#     projection <- data.frame(year = seq(2017, 2035+i, length = 20),
#                          countries = seq(100, 0, length = 20))
#     g <- g + geom_line(data = projection,
#             aes(x = year,
#                 y = countries,
#                 color = NA,
#                 group = NA),
#             color = 'grey',
#             alpha = 0.1) 
#     x <- g +
#   geom_line(data = projection,
#             aes(x = year,
#                 y = countries,
#                 color = NA,
#                 group = NA),
#             color = 'white',
#             lty = 2) 
#     print(x)
#     ggsave(filename = paste0('animation/',
#                              add_zero(counter, n = 3),
#                              '.png'))
#   }
# }
# make_animation()
```

Even in areas of high endemicity, advances in immunology, parasitology, modeling and vaccinology, along with rapid economic development, have made eradication appear a more feasible goal, though not possible in the short term [@Snow2015] [@Eckhoff2014]. From both administrative [@Yamey2004] and scientific [@Alonso2011] points of view, eradication has never before received so much attention, nor appeared so within grasp.

Most of the current research on expert opinion regarding the feasibility of malaria eradication focuses on the \emph{how} rather than the \emph{if} and the \emph{when} [@Tanner2015]. The participants in the Malaria Eradication Research Agenda (MalERA) process, in particular, have guided research goals and identified gaps in order for elmination to occur [@Alonso2011]. Though the MalERA authors firmly state that eradication is \emph{not} feasible given the "current tools and state of knowledge", mentions to the time frame are vague ("within the lifetime of young scientists just embarking on their careers") and no mention is made of the perception of the overall probability of achieving eradication.  Similarly, international programmes, such as the WHO Global Malaria Programme (GMP), have acknowledged the need "to take an official position on how and under what timeline malaria eradication could be achieved" [@WHO2015]. Clarity on timeline and likelihood of eradication could inform policy, and plays a crucial role in the economic analysis of the expected value of malaria elimination initiatives. But achieving clarity is difficult, given the many complex and interacting variables which affect malaria transmission, research funding, and technological development. The question of \emph{how} is complex enough, rendering questions of \emph{if} and \emph{when} even more difficult to answer.

The general objective of eradication serves to inspire, rally funder support, motivate researchers, and focus the efforts of public health practitioners. Proponents of disease eradication point to the success of historical and current campaigns (smallpox and polio, respectively), and highlight the benefits in health and wealth to future generations. However, the opportunity cost of investments in eradication-specific interventions can be high. And the “expected value” of an investment in a binary scenario (eradication or not) is a function of the probability of the scenario’s occurrence, and the temporal lag of that occurrence. Therefore, estimating the likelihood and time frame of eradication of malaria is essential for making sound investments in health. But how can we quantify these values given the great deal of uncertainty surrounding eradication?

Many studies have shown value in expert elicitation as a means to reduce uncertainty and inform decision making [@Morgan_2014]. But experts can sometimes be mistaken, a fact implicitly recognized when patients request a "second opinion" from another doctor when facing uncertainty. Additionally, relying on only a small panel of experts for a problem as complex as malaria exposes one to more bias, just as an overly concentrated investment portfolio exposes one to more risk. As Sir Francis Galton demonstrated in his famous study in which he showed that the crowds' aggregated estimates of cow's weight formed a quasi-normal distribution centered around the true weight [@Wallis2014] [@Galton1907Voxpopuli], averaging the perceptions of many can be more accurate than taking the opinion of any single expert, since the biases of diverse viewpoints can be complementary and symbiotic.

Assessing the likelihood and time frame of eradication, therefore, requires the combination of multiple points of view. Measuring consensus and discord among disease-specific researchers from a variety of disciplines can serve as a barometer of (informed) opinion, both guiding resources and identifying areas of concern [@Keenan2013]. We propose that the aggregation of malaria researchers' perceptions regarding the time frame andlikelihood of eradication forms a probability distribution which can be used to estimate the expected value of eradication-specific investments.

A great deal of previous research already covers the cost per case prevented [@Sicuri2011] [@Silumbe2015] [@Btto2016] [@IlungaIlunga2014] [@Dalaba2015]. Likewise, literature exists which could serve as a model for quantifying the location-specific opportunity costs associated with funneling funds towards malaria eradication [@Stuckey2014] [@White2011] [@Korenromp2012]. The correct discount rate for estimating the value of future lives saved is more of a philosophical question than an economic one. This leaves only the probability and time-frame to eradication, questions which have been addressed anecdotally, but never answered quantitatively.

The economic case for striving to achieve malaria eradication is compelling [@Barofsky2015]. Though the case-specific marginal cost of eradication can be expected to be high (relative to a simple control approach), successful eradication would mean massive recurring savings in the long-term [@Barrett2013]. However, to the extent that the case-specific marginal cost of prevention in an eradication campaign is high, estimating the likelihood of success is fundamental to the correct distribution of resources, particularly in low-income environments.

In other words, the rational assignment of resources for malaria eradication campaigns hinges on the expected value of those campaigns. We can describe this relationship below:

\begin{center}
$I_{T}$ = $P(x)_{}$ $\dfrac{B(m_{})}{m_{}}$(1 + $\delta$)$^{-T}$ -  $P(x)_{}$ $\dfrac{C(m_{})}{m_{}}$ (1 + $\delta$)$^{-T}$
\end{center}

\begin{itemize}
\item $ I $ is the return on investment
\item $ m $ is the number of malaria cases
\item $ x $ is whether eradication has been achievied
\item $ T $ is the time-frame (to both costs and benefits)
\item $ P $ is the probability (of eradication)
\item $ B $ is the benefit of preventing malaria
\item $ C $ is the cost of preventing malaria
\item $ \delta $ is the discount rate and opportunity cost
\end{itemize}

(I) is the return on investment at time (T) (the "end of malaria"). We take the present value of the benefits multiplied by the probability of succcess minus the value of costs times the probability of success, and multiplyboth terms by the discount rate traised to T to arrive at the return on investment.

The objective of this study is to help guide the optimal distribution of health resources by gauging (expert) opinion about, estimating the likelihood of, and quantifying the time frame to malaria eradication through a systemic survey of malaria research professionals from a wide array of academic disciplines.

# Methods

Our study population included all first authors (with available email addresses) returned in a PubMed search for the term "malaria" from January 1, 2010 through December 20, 2016. Personalized emails addressing the author by name and mentioning the relevant paper were sent to each of the `r nrow(email_df)` authors during the period from December 20, 2016 through January 2, 2017. Researchers were invited to participate by clicking a link to the survey form. The survey was simple, consisting of only name, email, and four content-related fields along with a "general comments" section. The survey was administered and data were collected through Google Drive. The original survey is viewable at [https://goo.gl/forms/IroAEooDuJ6KM5Ho2](https://goo.gl/forms/IroAEooDuJ6KM5Ho2).

Content-related survey fields consisted of:

1. Area of expertise.
2. Perceived probability (%) of malaria eradication in 10, 20, 30, 40, and 50 years.
3. Free choice perceived number of years until malaria eradication.
4. Ranking of the WHO's "health system building blocks" [@manya2017] in regards to attention needed in order to achieve eradication.

The purpose of final point was to introduce a prescriptive element to the researchers' responses (ie, identify potential areas of consensus in regards to the path towards eradication).

The survey was intentionally as short as possible, so as to appeal to time-pressed participants. However, supplementary data on researchers is of value for the assessing of selection bias and determinants of perception. Accordingly, we estimated the likelihood of a participant being of one of two genders (male or female) based on first name, and the likelihood of a participant being of one of five ethnicities (White, Black, Hispanic, Asian, Other) based on the provided last name. For gender estimations we used data from the North Atlantic Population Project, and U.S. government [@Lincoln]; for ethnicity, Khanna et al's bayesian prediction algorithm for the estimation of racial categories [@wru2015] [@Imai_2016] was employed.

The searching and retrieval of information from the PubMed database was carried out using the RISmed package [@Kovalchik2015].

Survey results were first analyzed descriptively. Following Francis Galton's example, we naively take the average of all responses as the likely point estimate for event probabilities, and the totality of the responses to each numeric question as the likely confidence interval of those likelihoods and time frames.

Quantitative analysis was carried out in R language [@R]. Ordinary least squares regression was employed to estimate the effect of area of expertise on the eradication pessimism-optimism (through the proxy of perceived years to eradication). De-biasing of sample selection was carried out using Heckman's 2 step approach as implemented by Toomet [@toomet2008]. All analysis code, as well as the code used for the identification and contacting of participants, is publicly available at [https://github.com/joebrew/malaria_survey](https://github.com/joebrew/malaria_survey).

Qualitative analysis of free text comments was carried out using thematic content analysis, with inductive open coding [@Markey_2014]. Participants were invited to provide "any general comments on the timeline and likelihood of global eradication".  Thematic content analysis [@Vaismoradi_2016] was employed to code responses following the 6-phase approach laid out by Braun and Clarke [@Braun2006]. The approach was inductive, used open-coding to identify latent themes, and underwent several iterations in which codes were modified, discarded and created. Using the RQDA software to assist in data management and theme coding [@Ronggui2017], four subject themes were identified. Comments were additionally coded as either descriptive (comments pertaining to the "problem" of malaria eradication) or presciptive (pertaining to potential "solutions" for eradicating malaria). Finally, free-text comments were scored for overall sentiment polarity [@sentimentr2017].

This research followed the principles of the Declaration of Helsinki. The study's protocol was approved by the Scientific Committee of the Barcelona Institute for Global Health.

# Results

```{r}
x <- df %>%
  filter(responded) %>%
  group_by(n_areas) %>%
  tally %>%
  mutate(p = n / sum(n, na.rm = TRUE) * 100)
```

A total of `r length(which(df$responded))` researchers participated in the survey from the `r nrow(df)` invitations sent (participation rate of `r round(length(which(df$responded)) / nrow(df) * 100, 2)`%). Areas of expertise were non-exclusive and self-described, with participants having the option to choose from up to 3 of 10 checkboxes, or to write in one or more "other" areas of expertise.\footnote{Response categories were Anthropology, Biochemistry, Bioinformatics, Biology, Chemistry, Clinical medicine, Drug discovery, Ecology, Economics, Entomology, Epidemiology, Geography, GIS, History, Immunology, Infectious disease, IT, Malaria, Medical entomology, Medicinal chemistry, Microbiology, Parasitology, Pharmacology, Pharmacy, Political science, Public health, Statistics, Vector biology, Vector control, and Virology.} `r sum(x$n[1:3])` (`r round(sum(x$p[1:3]), digits = 1)`%) participants declared three or fewer areas of expertise.

```{r}
temp <- areas_df %>%
  arrange(desc(count))
temp$area <- trimws(temp$area, which = 'both')
names(temp) <- c('Area', 'Researchers')
temp$Area <- as.character(temp$Area)
temp$Area <- Hmisc::capitalize(temp$Area)
temp$Area[temp$Area == 'It'] <- 'Information Technology'
# kable(temp[1:10,]) #%>%
  # kableExtra::kable_styling(#position = "float_right",
  #                           font_size = 7)
```

Participants had a total of `r length(old_areas)` unique (self-described) areas of expertise. The five most popular were `r paste0(paste0(temp$Area[1:5], ' (', temp$Researchers[1:5], ')'), collapse = ', ')`.

Participants were asked to quantify (as a percentage) the likelihood of eradication in the next 10, 20, 30, 40, and 50 years. Most participants saw eradication as extremely unlikely in the next 10-30 years, but increasingly likely thereafter. Figure 2 shows the distribution of year-specific likelihood perceptions (panels A-E), as well as an illustration of how both likelihood and uncertainty grow over time (panel F).

```{r, fig.cap='Perceptions of likelihood of eradication', fig.height = 4.5}
# Define function for analyzing the years_n columns
years_n <- function(n){
  x <- df %>%
    filter(responded) %>%
    dplyr::select_(paste0('years_', n)) 
  names(x) <- 'likelihood'
  result <- x %>%
    group_by(likelihood) %>%
    tally %>%
    mutate(p = n / sum(n) * 100) %>%
    filter(!is.na(likelihood))

  out <- list()
  out[[1]] <- result
  out[[2]] <-
    ggplot(data = result,
           aes(x = likelihood,
               y = p)) +
    geom_bar(stat = 'identity',
             fill = 'darkorange',
             alpha = 0.6) +
    labs(title = paste0(toupper(letters[n/10]),'. ', n, ' years'),
         # subtitle = 'Perceived likelihood of eradication',
         x = 'Likelihood',
         y = '%') +
    theme_paper(base_size = 10) +
    theme(axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(size=8))
  return(out)
}

# Make plots of all
years <- seq(10, 50, 10)
for (i in 1:length(years)){
  assign(paste0('g',i ),
         years_n(n = years[i])[[2]])
  x <- years_n(n = years[i])[[1]]
  x$year <- years[i]
  assign(paste0('x', i),
         x)

}
# Combine all data
x <- bind_rows(x1, x2, x3, x4, x5)

# Make an aggregated version
x$val <- ifelse(x$likelihood == '0%', 0,
                ifelse(x$likelihood == '1-20%', 10,
                       ifelse(x$likelihood == '21-40%', 30,
                              ifelse(x$likelihood == '41-60%', 50,
                                     ifelse(x$likelihood == '61-80%', 70,
                                            ifelse(x$likelihood == '81-99%', 90,
                                                   ifelse(x$likelihood == '100%', 100, NA)))))))
xa <- x %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(avg = weighted.mean(x = val, w = n, na.rm = TRUE),
            q3 = Hmisc::wtd.quantile(val, weights = n, 0.75, na.rm = TRUE),
            q1 = Hmisc::wtd.quantile(val, weights = n, 0.25, na.rm = TRUE))

g6 <- 
  ggplot(data = xa,
         aes(x = year,
             y = avg)) +
  geom_ribbon(aes(ymin = q1,
                  ymax = q3),
              fill = 'orange',
              alpha = 0.6) +
  geom_line(color = 'darkorange') +
  theme_paper(base_size = 10) +
  xlab('Years') +
  ylab('Average likelihood') +
  ggtitle('G. Aggregated perceptions',
          'Shaded area = inter-quartile range') +
  theme(plot.title = element_text(size=8))

Rmisc::multiplot(g1, g3, g5,
                 g2, g4, g6,
                 cols = 3)

# z <- ggplot(data = who %>% mutate(past = ifelse(past, 'Historical',
#                                            'WHO objective')) %>%
#          filter(!is.na(with_malaria)) ,
#        aes(x = year,
#            y = with_malaria,
#            color = past)) +
#   geom_line() +
#   geom_point() +
#   theme_paper() +
#   labs(x = '',
#        y = 'Countries',
#        title = 'Countries with malaria: Observed and WHO objectives',
#        subtitle = 'Data from http://www.who.int/malaria/areas/elimination/overview/en/') +
#   scale_color_manual(name = '',
#                      values = c('darkorange', 'blue')) +
#   ylim(0, 130) +
#   theme(axis.text.x = element_text(angle = 90)) +
#   geom_line(data = projection,
#             aes(x = year,
#                 y = countries,
#                 color = NA,
#                 group = NA),
#             color = 'white',
#             lty = 2) +
#   geom_hline(yintercept = 0, col = 'red')
# 
# # xa_plot <- xa %>% mutate(year = year + 2016) %>% filter(year == 2066)
# xa_plot <- data.frame(year = c(2015, 2066),
#                       val = c(100, 0))
#   
# z + geom_line(data = xa_plot,
#               aes(x = year,
#                   y = val),
#               color = 'white',
#               alpha = 0.6,
#               group = 1) +
#   theme_black()
```

Participants were asked to provide the number of years they believe it will take until eradication can be achieved. `r length(which(is.na(df$years[df$responded])))` (`r round(length(which(is.na(df$years[df$responded]))) / nrow(df) * 100, digits = 1)`%) were either blank or uninteligible, whereas `r length(which(!is.na(df$years)))` participants responded. Among respondants, `r length(which(df$years >= 50))` (`r round(length(which(df$years >= 50)) / length(which(!is.na(df$years))) * 100, digits = 1)` %) estimated that it would be 50 or more years until eradication.

```{r, fig.height = 3.5, fig.width = 4, eval = FALSE}
ggplot(data = df,
       aes(x = years)) +
  geom_histogram(fill = 'darkorange',
                 alpha = 0.6) +
  theme_paper(base_size = 10) +
  labs(x = 'Years',
       y = 'Responses',
       title = 'Perceived years until eradication',
       subtitle = '(Estimates of greater than 50 years reduced to 50)')
```

Respondents were qualitatively different from non-respondents. Importantly, the average number of total author-specific citations was `r round(mean(df$n_citations_total[df$responded]), digits = 2)` among respondents, but `r round(mean(df$n_citations_total[!df$responded]), digits = 2)` among non-respondents. This suggests either a tendency for more senior or impactful researchers not to respond. When examining the number of average citations per article, the difference between respondents remained: `r round(mean(df$n_citations_average[df$responded]), digits = 2)` among respondents, and `r round(mean(df$n_citations_average[!df$responded]), digits = 2)` among non-respondents, highlighting the greater impact of non-respondents relative to respondents. Males responded at a greater rate (`r round(mean(df$responded[df$sex == 'male'], na.rm = T) * 100, digits = 2)`) than females (`r round(mean(df$responded[df$sex == 'female'], na.rm = T) * 100, digits = 2)`) and those with Hispanic last names responded at a greater rate than those with last names of other ethnicities (see table).

```{r}
# function to get response rate per categorical variable
response_rate <- function(variable = 'sex'){
  df$x <- df[,variable]
  out <- df %>%
    group_by(x) %>%
    summarise(responded = length(which(responded)),
              n = n()) %>%
    ungroup %>%
    mutate(p = responded / n * 100)
  return(out)
}
out_list <- list()
vars <- c('sex',
            'ethnicity',
            'bin_citations')
for (i in 1:length(vars)){
  this_var <- vars[i]
  x <- response_rate(variable = this_var)
  x$variable <- this_var
  out_list[[i]] <- x
}
out <- bind_rows(out_list) %>%
  dplyr::select(variable, 
                x,
                responded,
                n, 
                p) %>%
  rename(Variable = variable,
         Characteristic = x,
         Responded = responded,
         Invited = n,
         `Response rate` = p)
out$Variable <- Hmisc::capitalize(out$Variable)
out$Variable[out$Variable == 'Bin_citations'] <- 'Citations'
out$Characteristic <- Hmisc::capitalize(out$Characteristic)
out$`Response rate` <- paste0(round(out$`Response rate`, digits = 2), ' %')

for(i in nrow(out):2){
  this_var <- out$Variable[i]
  prev_var <- out$Variable[i-1]
  if(this_var == prev_var){
    out$Variable[i] <- ''
  }
}
knitr::kable(out, 
             # format = 'latex', 
             caption = 'Response rate by author characteristics')
```

Selection bias is not of concern in the case of differential response if the groups for whom there are differences are not different in terms of the outcome variable. This was the case for sex: males responded at a significantly greater rate than females (p < 0.001), but were not statistically significantly different in regards to pessimism/optimism (ie, time frame or likelihood of eradication). In the case of researcher impact (as measured by the total number of citations), selection bias plays an important role: being highly-cited was associated both with eradication "pessimism" as well as likelihood of non-response. In other words, our pool of respondents was less highly-cited than our pool of invitees, and among respondents, those who were highly-cited tended to be more pessimistic (see Figure 3, Panel A).

\newpage

\noindent\fbox{%
    \parbox{\textwidth}{%
        Of the below two columns only one will remain in the paper: either (i) de-biasing using weighted binomial regression or (ii) de-biasing using Heckman.
    }%
}

\begin{multicols}{2}

\subsection*{De-biasing using weighted binomial regression}

In order to de-bias sample selection, we estimated a binomial logistic regression model for the likelihood of response as a function of sex and (binned) number of citations, and then used the inverse of the selection model's predictions as weights in a simple linear model to adjust our estimates. We run weighted model to estimate the likelihood of eradication at 10, 20, 30, 40, and 50 years (employing a separate model for each time period). Figure 3 shows the difference in the aggregate perceived likelihood of eradication before and after adjusting for sample selection.

```{r}
# Estimate model for de-biasing sample selection without Heckman.
response_model <- glm(responded ~ bin_citations + sex, data = df, family = binomial('logit'))
# Use model to predict likelihood of response
df$likelihood_of_response <- predict(response_model, df, type = 'response')
# Use inverse of likelihood of response to assign weight for model
df$weight <- 1 - df$likelihood_of_response
# Run likelihood of eradication models
for (i in seq(10, 50, 10)){
  assign(paste0('formula_simple_years_', i),
         as.formula(paste0('years_',
                           i,
                           '_numeric ~ sex + bin_citations')))
  assign(paste0('simple_', i),
         lm(get(paste0('formula_simple_years_', i)), data = df))
  assign(paste0('simple_', i, '_weighted'),
         lm(get(paste0('formula_simple_years_', i)), data = df,
            weights = df$weight))

}

# Get aggregate adjusted vs unadjusted likelihoods of eradications
simple <- data_frame(years = seq(10, 50, 10),
                     adjusted = NA,
                     unadjusted = NA)
for (i in seq(10, 50, 10)){
   simple$unadjusted[simple$years == i] <- mean(df[,paste0('years_', i, '_numeric')], na.rm = TRUE)
   simple$adjusted[simple$years == i] <- weighted.mean(x = df[,paste0('years_', i, '_numeric')],
                                                       w = ifelse(is.na(df$weight), 0,
                                                                  df$weight), na.rm = TRUE)
  # simple_model <- get(paste0('simple_', i))
  # weighted_model <- get(paste0('simple_', i, '_weighted'))

}
simple <- gather(simple, key, value, adjusted:unadjusted)
g2 <- ggplot(data = simple,
       aes(x = years,
           y = value,
           color = key,
           lty = key)) +
    geom_line(alpha = 0.8) +
  geom_point(alpha = 0.6) +
  theme_paper() +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'B: Perceived likelihood of eradication: Adjusted and unadjusted') +
  scale_linetype_manual(name = '', values = 1:2) +
  scale_color_manual(name = '', values = c('blue', 'darkorange'))
x <- df %>%
  group_by(citations = bin_citations) %>%
  summarise(years_10 = mean(years_10_numeric, na.rm = TRUE),
            years_20 = mean(years_20_numeric, na.rm = TRUE),
            years_30 = mean(years_30_numeric, na.rm = TRUE),
            years_40 = mean(years_40_numeric, na.rm = TRUE),
            years_50 = mean(years_50_numeric, na.rm = TRUE),
            n = length(unique(email))) %>%
  ungroup %>%
  gather(key, value, years_10:years_50) %>%
  filter(!is.na(citations))

cols <- colorRampPalette(c('red', 'orange', 'blue') #brewer.pal(n = 9, 'Spectral')
                         )(length(unique(x$citations)))
g1 <- ggplot(data = x %>% filter(!is.na(citations)),
       aes(x = as.numeric(gsub('years_', '', key)),
           y = value,
           color = citations,
           group = citations)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.6) +
  theme_paper() +
  scale_color_manual(name = 'Number of citations',
                     values = cols) +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'A: Difference in perceived likelihood of eradication by citations')
```

```{r, fig.height = 7.5}

Rmisc::multiplot(g1, g2, cols = 1)

# # Combine into a dataframe
# simple_df <- expand.grid(bin_citations = sort(unique(df$bin_citations)),
#                          sex = sort(unique(df$sex))) %>%
#   mutate(simple = NA,
#          weighted = NA)
# out_list <- list()
# counter <- 0
# for(i in seq(10, 50, 10)){
#   counter <- counter + 1
#   this_df <- simple_df %>%
#     mutate(years = i)
#   simple_model <- get(paste0('simple_', i))
#   weighted_model <- get(paste0('simple_', i, '_weighted'))
#   this_df$simple <- predict(simple_model, this_df)
#   this_df$weighted <- predict(weighted_model, this_df)
#   out_list[[counter]] <- this_df
# }
# simple_df <- bind_rows(out_list)
# simple_df <- simple_df %>%
#   gather(key, value, simple:weighted)
# # Just aggregate sexes
# simple_df <- simple_df %>%
#   group_by(bin_citations, years, key) %>%
#   summarise(value = mean(value)) %>%
#   ungroup
# 
# ggplot(data = simple_df,
#        aes(x = years,
#            y = value,
#            color = bin_citations)) +
#   geom_line()
```

\vfill\null
\columnbreak

\subsection*{De-biasing using Heckman}

In order to de-bias sample selection, we employed Heckman's two-step correction method. Given that sex affected response rate but not response itself, we were able to use sex as an instrumental variable, introducing it in both the selection and response models. Our selection models, which estimated the likelihood of survey response as a probit function of sex allowed us to estimate (via the Inverse Mills ratio) an expected error term for our outcome models, in which we quantify the effect of the number of citations on the perceived average likelihood of eradication at 10, 20, 30, 40, and 50 years (employing a separate model for each time period). Figure 3 shows the difference in the aggregate perceived likelihood of eradication before and after implementing Heckman's two-step model.

```{r}
x <- df %>%
  group_by(citations = bin_citations) %>%
  summarise(years_10 = mean(years_10_numeric, na.rm = TRUE),
            years_20 = mean(years_20_numeric, na.rm = TRUE),
            years_30 = mean(years_30_numeric, na.rm = TRUE),
            years_40 = mean(years_40_numeric, na.rm = TRUE),
            years_50 = mean(years_50_numeric, na.rm = TRUE),
            n = length(unique(email))) %>%
  ungroup %>%
  gather(key, value, years_10:years_50) %>%
  filter(!is.na(citations))

cols <- colorRampPalette(c('red', 'orange', 'blue') #brewer.pal(n = 9, 'Spectral')
                         )(length(unique(x$citations)))
x1 <- ggplot(data = x %>% filter(!is.na(citations)),
       aes(x = as.numeric(gsub('years_', '', key)),
           y = value,
           color = citations,
           group = citations)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.6) +
  theme_paper() +
  scale_color_manual(name = 'Number of citations',
                     values = cols) +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'A: Difference in perceived likelihood of eradication by citations')

x2 <- 
  ggplot(data = heckman,
       aes(x = years,
           y = value,
           color = key,
           lty = key)) +
  geom_line(alpha = 0.8) +
  geom_point(alpha = 0.6) +
  theme_paper() +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'B: Perceived likelihood of eradication: Heckman method and unadjusted') +
  scale_linetype_manual(name = '', values = 1:2) +
  scale_color_manual(name = '', values = c('blue', 'darkorange'))
```

```{r fig.height = 7.5}
Rmisc::multiplot(x1, x2, cols = 1)
```

```{r, eval = FALSE}
# Table 2 shows the Heckman model Heckman estimates for all five models.

rn <- names(heck_10$coefficients)
heck <- data_frame(Coefficient = rn,
                   `Model 10` = heck_10$coefficients,
                   `Model 20` = heck_20$coefficients,
                   `Model 30` = heck_30$coefficients,
                   `Model 40` = heck_40$coefficients,
                   `Model 50` = heck_50$coefficients)
for (j in 2:ncol(heck)){
  heck[,j] <- round(heck[,j], digits = 3)
}
heck$Coefficient <- gsub('bin_citations', 'Citations ', heck$Coefficient)
heck$Coefficient <- gsub('invMillsRatio', 'Inverse Mills Ratio', heck$Coefficient)
heck$Coefficient <- gsub('sigma', 'Sigma', heck$Coefficient)
heck$Coefficient <- gsub('rho', 'Rho', heck$Coefficient)
heck$Coefficient <- gsub('sexmale', 'Sex male', heck$Coefficient)
heck$Coefficient <- paste0(c(rep('Selection ', 4), rep('Outcome ', 2), rep('', 3)), heck$Coefficient)
knitr::kable(heck,
             caption = 'Heckman selection model coefficients')
```

\end{multicols}

\newpage

Participants were heterogenous in their perceptions, with all of the 5 pre-determined cut-offs containing the whole range of likelihoods (0-100%), and the free-form time frame responses suggesting that eradication would be achieved in a time-frame ranging from less than a decade to never. Heterogeneity can only be partly explained by area of expertise, as the differences in perception between disciplines were minor and insignificant (see Figure 4).

```{r, fig.height = 3, fig.width = 4, fig.cap = 'Average perceived years to aggregation by discipline'}
areas_df <- areas_df %>% arrange(desc(count))
small_areas <- areas_df$area[1:12]
model_formula <- paste0('years ~ ',
                        paste0(paste0('area_', gsub(' ', '_', small_areas)), collapse = ' + '))
model_formula <- as.formula(model_formula)
fit <- lm(model_formula, data = df)
# Get confidence intervals too
ci <- tidy(confint(fit))
x <- cbind(tidy(fit), ci)
names(x) <- c('term', 'estimate', 'se', 'stat', 'p', 'rownames', 'lwr', 'upr')
x <- x %>%
  dplyr::select(term, estimate,
                p, lwr, upr)
the_intercept <- x$estimate[1]
x <- x[-1,]
for (the_column in c('estimate', 'lwr', 'upr')){
  x[,the_column] <- x[,the_column] + the_intercept
}
x$term <- gsub('area_|TRUE', '', x$term)
x <- x %>%
  dplyr::arrange(estimate)
x$term <- gsub('_', '\n', x$term)
x$term <- Hmisc::capitalize(x$term)
x$term[x$term == 'It'] <- 'Information\ntechnology'

x$term <- factor(x$term, levels = x$term)
ggplot(data = x,
       aes(x = term,
           y = estimate)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  geom_errorbar(aes(ymin = lwr,
                    ymax = upr),
                alpha = 0.6,
                width = 0.8,
                color = 'red') +
  theme_paper(base_size = 10) +
  labs(x = 'Area',
       y = 'Years') +
  theme(axis.text.x = element_text(angle = 90))

```

```{r}
bb <- c('governance', 'health_workforce',
        'health_financing', 'health_technologies',
        'health_information', 'service_delivery')
results_list <- list()
for (i in 1:length(bb)){
  this_df <- data.frame(bb = bb[i],
                        ranking = as.numeric(df[,bb[i]]))
  results_list[[i]] <- this_df
}
results <- bind_rows(results_list)
results$bb <- gsub('_', '\n', results$bb)

results_agg <-
  results %>%
  dplyr::group_by(bb) %>%
  dplyr::summarise(avg = mean(ranking, na.rm = TRUE)) %>%
  arrange(avg)
results_agg$bb <- factor(results_agg$bb,
                         levels = results_agg$bb)
results$bb <- factor(results$bb, 
                     levels = levels(results_agg$bb))

# library(ggridges)
# ggplot(data = results %>% filter(!is.na(ranking)),
#        aes(x = ranking,
#            y = bb,
#            fill = bb)) +
#   geom_density_ridges(#fill = 'darkorange',
#                       alpha = 0.6,
#                       scale = 1.5) +
#   theme_paper() +
#   xlim(0,30) +
#   labs(x = 'Perceived importance',
#        y = 'WHO Building Block') +
#  scale_fill_cyclical(values = c("darkorange", "yellow"))

# ggplot(data = results) + 
#   geom_violin(aes(x = bb,
#            y = ranking),
#            fill = 'orange',
#               alpha = 0.6) +
#   geom_jitter(aes(x = bb,
#            y = ranking),
#            color = 'black',
#            with = 0.4,
#            height = 0.4,
#               alpha = 0.3,
#               size = 1) +
#   theme_paper(base_size = 10) +
#   labs(x = 'WHO Building block',
#        y = 'Ranking',
#        title = 'Perceived areas needing attention',
#        subtitle = 'Small points: jittered individual responses; large triangles: averages') +
#   geom_point(data = results_agg,
#              aes(x = bb,
#                  y = avg),
#              size = 5,
#              pch = 17,
#              color = 'black',
#              alpha = 0.8) +
#   ylim(0, 6)

# a <- data_frame(category = rep(c("Innovation",
#                                                    "Systemic Challenges",
#                                                    "Poverty",
#                                                    "Meta"),2),
#                 type = rep(c('Prescriptive', 'Descriptive'), each = 4),
#                 val = c(213,
#                         25,
#                         85,
#                         32,22,
#                         112,
#                         70,
#                         142))
# a$val <- ifelse(a$type == 'Prescriptive', a$val * -1, a$val)
# ggplot(data = a,
#        aes(x = category,
#            y = val)) +
#   geom_bar(alpha = 0.5,stat = 'identity', aes(fill = type)) +
#   coord_flip() +
#   geom_label(aes(label = abs(val))) +
#   theme(legend.position = 'bottom') +
#   scale_fill_manual(name = '',
#                     values = c('lightblue', 'darkorange')) +
#   theme_fivethirtyeight() +
#   ylim(-220, 150) +
#   labs(title = 'Classification of survey free responses',
#        subtitle = 'By thematic categorization and diagnostic')
```

## Qualitative analysis of comments

Of the `r length(which(df$responded))` who responded to the survey, `r length(which(!is.na(df$comments)))` (`r round(length(which(!is.na(df$comments))) / length(which(df$responded)) * 100, digits = 2)`%) provided a comment. Relative to non-commenters, commenters were slightly more optimistic on average, but also slightly more polarized in opinion. The four subject themes identified through iterative, open coding were:

1. *Innovation*: comments pertaining to new tools, drugs, methods, or strategies required to achieve eradication.
2. *Systemic challenges*: comments pertaining to political, social, environmental or logistical issues related to eradication.
3. *Complexity*: comments which of focus on the multi-dimensional components of eradication.
4. *Tackling the fundamentals*: challenging the question of eradication itself, re-framing the goal, insisting on other priorities, etc.

Comments were also classified as descriptive or prescriptive. A slight majority (`r round(320 / 540 * 100, digits = 2)`%) were descriptive. Descriptive commenters were slightly more pessimistic (in regards to perceived years until eradication) than prescriptive commenters, though this difference did not reach the level of statistical significance (p = 0.21). Descriptive comments also received sentiment polarity scores which were slightly more negative than prescriptive comments, although again this difference did not reach the level of statistical significance (p = 0.18).

In regards to **innovation**, comments largely pertained to the necessity of further technological advances. One participant wrote that "currently available technology can't achieve [eradication], even if delivered optimally"; another argued that eradication could not be achieved without a "game-changing innovation", whereas multiple others referred to the need for "transformative" technologies:

\begin{quote}
We can't achieve eradication with our current tools. We'd need new medicines, a better vaccine, and maybe other vector control tools.
\end{quote}

Many noted the need to "overcome the challenge of drug resistance". More than 10% of commenters noted the need for an effective vaccine. Genetic engineering was mentioned by several commenters as a promising means to achieve eradication quickly. Most comments coded as "innovation" were presciptive in nature, often suggesting the nature of the needed innovation, with a heavy slant towards pharmaceutical options and vaccination.

**Systemic challenges** to malaria eradication were noted by the majority of commenters. Comments in this category can be divided into roughly four sub-themes: (i) lack of coordination, (ii) lack of good surveillance and health services delivery, (iii) lack of political will and (iv) poverty. In direct contrast to the previous comments, many emphasized that "we already have the tools to achieve  eradication" and that the only piece lacking was "robust health systems". Many commenters noted problems of coordination, as illustrated in the below quote.

\begin{quote}
It will be very difficult to eradicate malaria... not because we don't have the technologies, which we already have... the problem is politics. Malaria doesn't stop in (sic) borders of a country and it would take a joint effort of a lot of political leaders to get programs in place to fight malaria. Unfortunately I don't see this happening anytime soon.
\end{quote}

Others echoed the sentiment, with many comments focusing on the need of strong surveillance and treatment delivery systems. Many commenters focused on other reasons for stagnating progress, such as "weak or failing health systems...due to political unwillingness or conflict".  Comments coded as "systemic" tended to be descriptive and slightly more pessimistic than others. Many noted that malaria is a "disease of poverty", with "social injustice" as the root cause. Some made the sequential argument that "eradication of poverty" must preceed disease eradication. Along the same lines, one commenter wrote:

\begin{quote}
Eradication requires a full systems-wide approach, not a disease-specific approach. The eradication of smallpox was a triumph of management, not medicine or technology.
\end{quote}

Another noted that the survey "left off the list the most important factor - economic development". Many echoed the sentiment, stating that "with no economic development, you cannot have eradication" and that poverty is the "cause" of malaria. As with the "systemic challenges" of eradication, comments which were coded as poverty tended to be descriptive rather than prescriptive.

**Complexity** was a relatively rare category (<20\% of all comments), those whose comments were coded as the "complexity" category were slightly more pessimistic than average in regards to the timeline and likelihood of eradication. Many commenters highlighted the inherent challenges in the epidemiology of malaria, such as the changing dynamics of malaria transmission, the resilience of the parasite and vector, climate change, and the inability to aim interventions accurately with an "ever-moving target". The potential for adaptation was highlighted in reference both to the mosquito as well as the parasite itself. Many comments addressed the fact that the conversation on eradication is largely taking place within the public health community, whereas the causes of malaria endemicity are largely orthogonal to public health interventions, "going beyond the health sector". Several commenters pointed out the multitude of prerequisite conditions for eradication to even be considered feasible:

\begin{quote}
To my mind, this question is highly dependent on background context, e.g. global political and economic dynamics, as well as international conflict. Complete global eradication is an extremely singular goal that requires a vast array of necessary conditions - if any of these fail, eradication will not be achieved.  
\end{quote}

**Tackling the fundamentals** was rarer than the other categories, but cohesive and consistent. One commenter wrote that eradication is a "postwar" idea that developed from the "abandonment of a broad sociopolitical understanding of the causes of disease, and the emphasis on technological solutions." Many stated that global malaria eradication was simply not possible, and two argued that it may not be desirable or ethical. One stated that the concept was "absurd" and that "I'm not even sure why people talk about it". Another questioned the utility of discussing "eradiction" as a concept:

\begin{quote}
Eradication is a different objective than elimination. Elimination means that the disease is not endemic but could reappear even in a country like Norway if infrastructure breaks down. Elimination may be possible in poor endemic countries, following socioeconomic development. Eradication means that the parasite disapears from the planet, which is not realistic...
\end{quote}

Many pointed to the failure of the WHO GMP in the 1950s as evidence of the futile nature of eradication programs. Many highlighted that talking about "malaria" as one disease misses the mark, since the different species of parasite and contexts in which they live make elimination in each area very distinct from other areas. "Meta" comments tended to be skeptical of the feasibility of eradication, or to challenge the  notion that it is a meaningful or useful endeavor. Largely, they were prescriptive, advocating for a re-framing of the conversation so that the focus was not on an "arbitrary" goal like eradication, but rather on scaling up control and making region-specific progress.

# Discussion

Our survey reveals a notable gap between the public discourse on malaria eradication, and those views held (largely privately) by malaria researchers. Approximately three-quarters of respondents believe that malaria will not be eradicated in the next 50 years; in other words, assuming our pool of respondents is of typical post-PhD age, most believe they will not live to see eradication.

If we take the pooled opinions of researchers, as revealed through this survey, as the best approximation of truth available, then the disconnect between public discourse on malaria eradication and private opinion is troubling. It suggests that the goals set by the Bill and Melinda Gates Foundation (eradication "within a generation") and the WHO GMP (eradication "within the lifetime of young scientists just embarking on their careers") are unrealistic at best, and misleading at worst. This has important implications for spending in public health, since the attention and resources of funders, researchers and government agencies are directed to those areas where a result is expected.

## Limitations

There are many reasons to be skeptical that the combined "wisdom of crowds" of malaria researchers is a reliable indicator of truth. First, academic researchers are specialists - their narrow, field-specific view of eradication's feasibility is arguably less reliable than the more "global" views of the WHO, or Bill Gates. Second, a number of biases may come into play in this study.

Among these biases, four are worth mentioning specifically. (1) "Conjunction fallacy" suggests that the general goal of eradication may seem less likely than the sum of the goals of country-specific elimination. (2) A (reverse) variant of the "hot hand fallacy", in which researchers may mistakenly base their assessment of current chances of eradication on previous failures. (3) Parkinson's law of triviality suggests that researchers may disproportionately see the challenges of their own research (antimalarial drug resistance, etc.) as larger or more relevant to the global eradication campaign than they really are. (4) Finally, and ironically, "optimism bias" may play a perverse role in researchers' responses; though eradication is certainly a goal desired by all, one could argue that malaria research specialists subconciously realize that they actually stand to lose out profesionally in the case of eradication.

We made no distinguishment between quality of research, years of experience, academic achievements, geographical area, etc. Our inclusion criteria were simple and rudimentary, allowing us to access many diverse viewpoints, but (perhaps inappropriately) weighting all viewpoints as equal.

This study included the first authors of indexed journals. Though certainly a group with important knowledge related to malaria, this misses malaria control program employees, health agency workers, and other stakeholders. Their experiences and viewpoints may be different from those of academics, and arguably more relevant. Additionally, in public health journals, first authors are often more junior than their collaborators. By focusing solely on first authors, we may have unintentionally created a sample which is qualitatively different than the true universe of malaria researchers.

## Conclusion

Whether the collective "wisdom" of researchers is worth adhering to or not is arguable. That said, to the extent that researchers' private opinions differ so greatly from those of global health leaders is itself a noteworthy result. If leaders' high levels of optimism are well-placed, this suggests that researchers are unduly pessimistic, and efforts should be made to convince them that eradication is more feasible than they realize (so that efforts are directed accordingly). If, on the other hand, collective wisdom is a better forecasting agent than the states objectives of global health leaders, then it is important to recognize the sad truth that we are likely further from eradication than we would like to be. In this case, though the discourse on eradication can be motivating and inspiring, resources should be directed in a manner that reflects an intended outcome's likelihood of occurrence.

Our survey attempted to quantify the likelihood of and time frame to malaria eradication by gauging the collective opinion of malaria researchers. Our results suggest that eradication is unlikely in the short and medium terms, and that there exists a great deal of disagreement among researchers in regards to which areas require the most attention in order for eradication to occur. The causes for pessimism are diverse, but common themes were the need for innovation, systemic challenges, and the complexity of the disease and its transmission, along with a notable share of researchers who question the utility of the eradication concept itself. The implication of these results are three-fold: (1) that those working or investing in eradication-specific campaigns should factor in the low likelihood and long time frame when calculating those campaigns' expected value; (2) that champions of near-term eradication need to make a more compelling case to malaria researchers of eradication's feasibility, in order to better focus and inspire the latter; and (3) that discord among malaria researchers as to which areas need the most attention in order for eradication to occur, suggests a need for open discussion and concensus-making, so that resources and focus are directed appropriately.

References {#references .unnumbered}
==========