### 2018-03-22 paper.Rmd
major update

---
title: "Researchers' perceptions of malaria eradication: findings from a mixed-methods analysis of a large online survey"
bibliography: library.bib
output: 
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
csl: elsevier-with-titles.csl
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
# No scientific notation
# options(scipen=999)

# Packages (load child-specific packages in their respective files)
library(knitr)
library(xtable)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
               fig.align = 'center',
               fig.pos = 'h')
options(xtable.comment = FALSE)

# Get data
source('prepare_data.R', echo = FALSE)

theme_paper <- theme_publication
```

```{r}
authors <- data.frame(name = c('Joe Brew',
                               'Menno Pradhan',
                               'Jacqueline Broerse',
                               'Elisa Sicuri'),
                      email = c('joe@databrew.cc',
                                'm.p.pradhan@vu.nl',
                                'j.e.w.broerse@vu.nl',
                                'elisa.sicuri@isglobal.org'),
                      affilitation = c("isglobal,cism,vu",
                                       "vu,uva",
                                       "vu",
                                       "isglobal,icl,cism"),
                      footnote = c("Corresponding Author",
                                   rep('', 3)),
                      stringsAsFactors = FALSE)
institutions <- data.frame(code = c('isglobal',
                                    'icl',
                                    'cism',
                                    'vu',
                                    'uva'),
                           institution = c('Barcelona Institute for Global Health',
                                           'Imperial College London',
                                           'Centro de Investigação em Saúde de Manhiça',
                                           'VU University Amsterdam',
                                           'University of Amsterdam'),
                           address = c('c/ Rosselló, 132, 5è 2a. 08036, Barcelona, Spain',
                                       'South Kensington Campus, London SW7 2AZ, U.K.',
                                       'Vila da Manhiça, Bairro Cambeve, Rua 12, Distrito da Manhiça, CP 1929, Maputo, Mozambique',
                                       'De Boelelaan 1105, 1081 HV Amsterdam, Netherlands',
                                       'REC E, Roetersstraat 11, Amsterdam, Netherlands'),
                      stringsAsFactors = FALSE)

make_author <- function(authors,
                        institutions){

  out <- c()
  for (i in 1:nrow(authors)){
    this_author <- authors$name[i]
    these_institutions <- authors$affilitation[i]
    these_institutions <- unlist(strsplit(these_institutions, ','))
    author_text <- paste0(this_author)
    for (j in 1:length(these_institutions)){
      this_institution <- these_institutions[j]
      if(any(grepl(this_institution, out))){
        author_text <- paste0(author_text,
                              '\\footrecall{',
                              this_institution,
                              '}'
                              )
      } else {
        # new institution, get full name
        full_name <- institutions$institution[institutions$code == this_institution]
        address <- institutions$address[institutions$code == this_institution]
        author_text <- paste0(author_text,
                              '\\footremember{',
                              this_institution,
                              '}{',
                              full_name,
                              ': ',
                              address,
                              '}')
      }
    }
    out[i] <- author_text
  }
  out <- paste0(out, collapse = ', ')
  cat(out)
}
```

\begin{center}
\begin{large}

```{r, results = 'asis', eval = TRUE}
make_author(authors = authors,
            institutions = institutions)
```

\end{large}
\end{center}

\vspace{5mm}

\begin{center}
\textbf{Abstract}  
\end{center}

\vspace{5mm}
\begin{center}
\begin{changemargin}{3cm}{3cm}

Quantifying an event's probability and time frame is essential for calculating its expected value. In the case of global malaria eradication, uncertainty regarding feasibility makes it difficult for policymakers and public health practitioners to make fully informed decisions. The opportunity cost of investments in eradication-specific interventions can be high, particularly in contexts with other urgent health priorities. In a large online survey of malaria researchers (n = 844), we query perceptions regarding the likelihood and time frame of eradication. We assess pessimism/optimism (via the proxy of years-to-eradication), broken down by area of expertise, and adjust for selection bias. We also carry out a qualitative analysis of free-form comments provided by 61\% of respondents. Our results show that researchers are largely sceptical of the feasibility of near-term malaria eradication, with most believing that eradication will not be achieved for 50 years or more.
\end{changemargin}
\end{center}

\noindent\fbox{%
    \parbox{\textwidth}{%
        \subsection*{Research Highlights}
        \begin{itemize}
          \item Makes novel use of a "wisdom of crowds" approach
          \item Assesses perceived time frame and likelihood of global malaria eradication among researchers
          \item Identifies challenges pertaining to eradication through qualitative analysis
          \item Highlights the gap between researchers' collective pessimism and institutional optimism
        \end{itemize}
      \subsection*{Keywords}
        Malaria, eradication, elimination, mixed methods, survey, crowdsourcing 
    }%
}
m

# Introduction

## Background

Malaria is a parasitic disease transmitted among humans by mosquitoes. The _Plasmodium falciparum_ species accounts for a large majority of the 200 million annual cases as well as the half million annual deaths worldwide [@White2014] [@World2016]. Malaria "elimination", the "interruption of all local transmission of the infection in a country or region" [@Tanner2015] is actively being pursued by dozens of countries around the world, leading to a renewed push for "eradication" (the global elmination of malaria) [@Rabinovich_2017].

This is not the first time that eradication has been in the international spotlight. The scientific and public health communities have had eradication on their policy agenda since the WHO established the Global Malaria Eradication Program in the 1950s [@Njera2011]. In 1957, U.S. President Dwight Eisenhower told Congress that malaria could be expected to be eradicated in five years. Following the failure of the WHO's first attempt, the focus shifted away from global eradication towards local elimination and control strategies.  In recent years, much of the discourse regarding malaria has shifted back to global eradication [@Roberts2007], with funders, researchers, and public health practitioners rallying to the cause [@Tanner2015]. The Bill and Melinda Gates Foundation has begun actively promoting eradication as feasible "within a generation" [@Gates2014], and the World Health Organization (WHO) has set ambitious goals, stating the objective of eliminating malaria in 65 new endemic countries from 2015 through 2030 (Figure 1).

```{r, fig.cap='Countries with malaria: Observed and WHO objectives', fig.height = 3, fig.width = 4}
# Read WHO data (compiled manually from WHO website)
# Data from http://www.who.int/malaria/areas/elimination/overview/en/'
who <- readr::read_csv('who_elimination_data.csv')
who$year <- as.numeric(substr(who$year, 6, 9))
projection <- data.frame(year = seq(2030.2, 2035, length = 20),
                         countries = seq(33, 0, length = 20))
g <- ggplot(data = who %>% mutate(past = ifelse(past, 'Historical',
                                           'WHO objective')) %>%
         filter(!is.na(with_malaria)) ,
       aes(x = year,
           y = with_malaria,
           color = past)) +
  geom_line() +
  geom_point() +
  theme_paper() +
  labs(x = '',
       y = 'Countries') +
  scale_color_manual(name = '',
                     values = c('darkorange', 'blue')) +
  ylim(0, 130) +
  theme(axis.text.x = element_text(angle = 90)) #+
  # geom_line(data = projection,
  #           aes(x = year,
  #               y = countries,
  #               color = NA,
  #               group = NA),
  #           color = 'black',
  #           lty = 2) +
  # geom_hline(yintercept = 0, col = 'red') 
g
# #Make animation for presentation
# make_animation <- function(){
#   g <- 
#     ggplot(data = who %>% mutate(past = ifelse(past, 'Historical',
#                                            'WHO objective')) %>%
#          filter(!is.na(with_malaria)) ,
#        aes(x = year,
#            y = with_malaria,
#            color = past)) +
#   geom_line() +
#   geom_point() +
#   theme_black() +
#   labs(x = '',
#        y = 'Countries',
#        title = 'Countries with malaria: Observed and WHO objectives',
#        subtitle = 'Data from http://www.who.int/malaria/areas/elimination/overview/en/') +
#   scale_color_manual(name = '',
#                      values = c('darkorange', 'blue')) +
#   ylim(0, 130) +
#   theme(axis.text.x = element_text(angle = 90)) +
#   geom_hline(yintercept = 0, col = 'red') +
#     xlim(1970, 2060)
#   add_zero <- function (x, n) {
#     x <- as.character(x)
#     adders <- n - nchar(x)
#     adders <- ifelse(adders < 0, 0, adders)
#     for (i in 1:length(x)) {
#         if (!is.na(x[i])) {
#             x[i] <- paste0(paste0(rep("0", adders[i]), collapse = ""), 
#                 x[i], collapse = "")
#         }
#     }
#     return(x)
# }
#   counter <- 0
#   a <- (seq(1,5, length = 15))^2.5
#   a <- round(a)
#   a <- c(a, rev(a) +2)
#   for (i in a){
#     message(counter)
#     counter <- counter +1
#     projection <- data.frame(year = seq(2017, 2035+i, length = 20),
#                          countries = seq(100, 0, length = 20))
#     g <- g + geom_line(data = projection,
#             aes(x = year,
#                 y = countries,
#                 color = NA,
#                 group = NA),
#             color = 'grey',
#             alpha = 0.1) 
#     x <- g +
#   geom_line(data = projection,
#             aes(x = year,
#                 y = countries,
#                 color = NA,
#                 group = NA),
#             color = 'white',
#             lty = 2) 
#     print(x)
#     ggsave(filename = paste0('animation/',
#                              add_zero(counter, n = 3),
#                              '.png'))
#   }
# }
# make_animation()
```

Even in areas of high endemicity, advances in immunology, parasitology, modeling and vaccinology, along with rapid economic development, have made eradication appear a more feasible goal, though not possible in the short term [@Snow2015] [@Eckhoff2014]. From both administrative [@Yamey2004] and scientific [@Alonso2011] points of view, eradication has never before received so much attention.

Most of the current research on expert opinion regarding the feasibility of malaria eradication focuses on the \emph{how} rather than the \emph{if} and the \emph{when} [@Tanner2015]. The participants in the Malaria Eradication Research Agenda (MalERA) process, in particular, have guided research goals and identified gaps in order for elmination to occur [@Alonso2011] [@Rabinovich_2017]. Though the MalERA authors firmly state that eradication is \emph{not} feasible given the "current tools and state of knowledge", mentions to the time frame are vague ("within the lifetime of young scientists just embarking on their careers" [@Alonso2011] and "in the longer term" [@Rabinovich_2017]) and no mention is made of the perception of the overall probability of achieving eradication.  International programmes, such as the WHO Global Malaria Programme (GMP), have acknowledged the need "to take an official position on how and under what timeline malaria eradication could be achieved" [@WHO2015]. Clarity on timeline and likelihood of eradication could inform policy, and plays a crucial role in the economic analysis of the expected value of malaria elimination initiatives. But achieving clarity is difficult, given the many complex and interacting variables which affect malaria transmission, research funding, and technological development. The question of \emph{how} is complex enough, rendering questions of \emph{if} and \emph{when} even more difficult to answer.

The general objective of eradication serves to inspire, rally funder support, motivate researchers, and focus the efforts of public health practitioners. Proponents of disease eradication point to the success of historical and current campaigns (smallpox and polio, respectively), and highlight the long term benefits in health and wealth to future generations (one often cited figure for smallpox was a cost-benefit ratio of more than 400 to 1 [@Barrett_2006]). However, the opportunity cost of investments in eradication-specific interventions is likely to be high, particularly in contexts of very low resources and many competing health problems. Additionally the impact of malaria on economic indicators is heterogeneous, with singificant impact in areas of Sub-Saharan Africa - where the endemicity is high - and little to no impact in areas of low endemicity [@Cutler_2010] [@Barofsky2015]. Additionally, the “expected value” of an investment in a binary scenario (eradication or not) is a function of the probability of the scenario’s occurrence, the temporal lag of that occurrence, and the uncertainy around both probability of occurrence and temporal lag. This uncertainty makes the "risk" of investing in malaria elimination and eradication high. Therefore, estimating the likelihood and time frame of eradication of malaria is essential for making sound investments in health. But how can we quantify these values given the great deal of uncertainty surrounding eradication?

Many studies have shown value in expert elicitation as a means to reduce uncertainty and inform decision making [@Morgan_2014]. But experts can sometimes be mistaken, a fact implicitly recognized when patients request a "second opinion" from another doctor when facing uncertainty. Additionally, relying on only a small panel of experts for a problem as complex as malaria exposes one to bias, just as an overly concentrated investment portfolio exposes one to more risk. As Sir Francis Galton demonstrated in his famous study in which he showed that the crowds' aggregated estimates of cow's weight formed a quasi-normal distribution centered around the true weight [@Wallis2014] [@Galton1907Voxpopuli], averaging the perceptions of many can be more accurate than taking the opinion of any single expert, since the biases of diverse viewpoints can be complementary and symbiotic.

Assessing the likelihood and time frame of eradication, therefore, requires the combination of multiple points of view. Measuring consensus and discord among disease-specific researchers from a variety of disciplines can serve as a barometer of (informed) opinion, both guiding resources and identifying areas of concern [@Keenan2013]. We propose that the aggregation of malaria researchers' perceptions regarding the time frame andlikelihood of eradication forms a probability distribution which can be used to estimate the expected value of eradication-specific investments.

A great deal of previous research from serveral endemic areas already covers the cost per case prevented [@Sicuri2011] [@Silumbe2015] [@Btto2016] [@IlungaIlunga2014] [@Dalaba2015]. Likewise, literature exists which could serve as a model for quantifying the location-specific opportunity costs associated with funneling funds towards malaria eradication [@Stuckey2014] [@White2011] [@Korenromp2012]. The correct discount rate for estimating the value of future lives saved is more of a philosophical question than an economic one. This leaves only the probability and time-frame to eradication, questions which have been addressed anecdotally, but never answered quantitatively.

Though the case-specific marginal cost of eradication can be expected to be high (relative to a simple control approach), successful eradication would mean massive recurring savings in the long-term [@Barrett2013]. However, to the extent that the case-specific marginal cost of prevention in an eradication campaign is high, estimating the likelihood of success is fundamental to the correct distribution of resources, particularly in low-income environments.

In other words, the rational assignment of resources for malaria eradication campaigns hinges on the expected value of those campaigns. We can describe this relationship below:

\begin{center}
$I_{T}$ = $P(x)_{}$ $\dfrac{B(m_{}*p)}{m_{}*p}$(1 + $\delta$)$^{-T}$ -  $P(x)_{}$ $\dfrac{C(m_{}*p)}{m_{}*p}$ (1 + $\delta$)$^{-T}$
\end{center}

\begin{itemize}
\item $ I $ is the return on investment
\item $ m $ is the prevalence of malaria
\item $ p $ is the population size
\item $ x $ is whether eradication has been achievied
\item $ T $ is the time-frame (to both costs and benefits)
\item $ P $ is the probability (of eradication)
\item $ B $ is the benefit of preventing malaria
\item $ C $ is the cost of preventing malaria
\item $ \delta $ represents both the discount rate and opportunity cost
\end{itemize}

(I) is the return on investment at time (T) (the "end of malaria"). We take the present value of the benefits multiplied by the probability of succcess minus the value of costs times the probability of success, and multiplyboth terms by the discount rate traised to T to arrive at the return on investment.

The objective of this study is to gauge (expert) opinion about, estimate the likelihood of, and quantify the potential time frame to malaria eradication through a systematic online survey of malaria research professionals from a wide array of academic disciplines. In doing so, we aim to help guide the optimal distribution of health resources by informing estimations of the expected value of malaria eradication efforts.

# Methods

Our study population included all first authors (with available email addresses) returned in a PubMed search for the term "malaria" from January 1, 2010 through December 20, 2016. We used PubMed because it was the most comprehensive publication database for malaria-releated research, and also exposed enough metadata about articles and authors so as to allow for relevant analysis. Personalized emails addressing the author by name and mentioning the relevant paper were sent to each of the `r nrow(email_df)` authors during the period from December 20, 2016 through January 2, 2017. Researchers were invited to participate by clicking a link to the survey form. The survey was simple, consisting of only name, email, and four content-related fields along with a "general comments" section. The survey was administered and data were collected through Google Drive. The original survey is viewable at [https://goo.gl/forms/IroAEooDuJ6KM5Ho2](https://goo.gl/forms/IroAEooDuJ6KM5Ho2).

Content-related survey fields consisted of:

1. Area of expertise.
2. Perceived probability (%) of malaria eradication in 10, 20, 30, 40, and 50 years.
3. Free choice perceived number of years until malaria eradication.

The survey was intentionally as short as possible, so as to appeal to time-pressed participants. However, supplementary data on researchers is of value for the assessing of selection bias and determinants of perception. Accordingly, we estimated the likelihood of a participant being of one of two genders (male or female) based on first name. In order to do this, we used data from the North Atlantic Population Project, and U.S. government [@Lincoln]. We also calculated the total number of citations which each author had received, as well as the total number of articles authored. We binned total citations into three categories: 0-5 (junior-level researchers, PhD students, etc.); 6-99 (most professional academics); and >99 (the most prolific researchers). The searching and retrieval of information pertaining to articles and citations from the PubMed database was carried out using the RISmed package [@Kovalchik2015].

Survey results were first analyzed descriptively. Following Francis Galton's example, we naively take the average of all responses as the likely point estimate for event probabilities, and the totality of the responses to each numeric question as the likely confidence interval of those likelihoods and time frames.

Quantitative analysis was carried out in R language [@R]. Ordinary least squares regression was employed to estimate the effect of area of expertise on perceived years to eradication. All analysis code, as well as the code used for the identification and contacting of participants, is publicly available at [https://github.com/joebrew/malaria_survey](https://github.com/joebrew/malaria_survey).

Qualitative analysis of free text comments was carried out using thematic content analysis, with inductive open coding [@Markey_2014]. Participants were invited to provide "any general comments on the timeline and likelihood of global eradication".  Thematic content analysis [@Vaismoradi_2016] was employed to code responses following the 6-phase approach laid out by Braun and Clarke [@Braun2006]. The approach was inductive, used open-coding to identify latent themes, and underwent several iterations in which codes were modified, discarded and created. Using the RQDA software to assist in data management and theme coding [@Ronggui2017], four subject themes were identified. Comments were additionally coded as either descriptive (comments pertaining to the "problem" of malaria eradication) or presciptive (pertaining to potential "solutions" for eradicating malaria). Finally, free-text comments were scored for overall sentiment polarity [@sentimentr2017].

# Results

```{r, fig.cap='Perceptions of likelihood of eradication', fig.height = 4}
# Define function for analyzing the years_n columns
years_n <- function(n){
  x <- df %>%
    filter(responded) %>%
    dplyr::select_(paste0('years_', n)) 
  names(x) <- 'likelihood'
  result <- x %>%
    group_by(likelihood) %>%
    tally %>%
    mutate(p = n / sum(n) * 100) %>%
    filter(!is.na(likelihood))

  out <- list()
  out[[1]] <- result
  out[[2]] <-
    ggplot(data = result,
           aes(x = likelihood,
               y = p)) +
    geom_bar(stat = 'identity',
             fill = 'darkorange',
             alpha = 0.6) +
    labs(title = paste0(toupper(letters[n/10]),'. ', n, ' years'),
         # subtitle = 'Perceived likelihood of eradication',
         x = 'Likelihood',
         y = '%') +
    theme_paper(base_size = 10) +
    theme(axis.text.x = element_text(angle = 90)) +
  theme(plot.title = element_text(size=8)) +
    ylim(0, 60)
  return(out)
}

# Make plots of all
years <- seq(10, 50, 10)
for (i in 1:length(years)){
  assign(paste0('g',i ),
         years_n(n = years[i])[[2]])
  x <- years_n(n = years[i])[[1]]
  x$year <- years[i]
  assign(paste0('x', i),
         x)

}
# Combine all data
x <- bind_rows(x1, x2, x3, x4, x5)

# Make an aggregated version
x$val <- ifelse(x$likelihood == '0%', 0,
                ifelse(x$likelihood == '1-20%', 10,
                       ifelse(x$likelihood == '21-40%', 30,
                              ifelse(x$likelihood == '41-60%', 50,
                                     ifelse(x$likelihood == '61-80%', 70,
                                            ifelse(x$likelihood == '81-99%', 90,
                                                   ifelse(x$likelihood == '100%', 100, NA)))))))
xa <- x %>%
  dplyr::group_by(year) %>%
  dplyr::summarise(avg = weighted.mean(x = val, w = n, na.rm = TRUE),
            q3 = Hmisc::wtd.quantile(val, weights = n, 0.75, na.rm = TRUE),
            q1 = Hmisc::wtd.quantile(val, weights = n, 0.25, na.rm = TRUE))

g6 <- 
  ggplot(data = xa,
         aes(x = year,
             y = avg)) +
  geom_ribbon(aes(ymin = q1,
                  ymax = q3),
              fill = 'orange',
              alpha = 0.6) +
  geom_line(color = 'darkorange') +
  theme_paper(base_size = 10) +
  xlab('Years') +
  ylab('Average likelihood') +
  ggtitle('G. Aggregated perceptions',
          'Shaded area = inter-quartile range') +
  theme(plot.title = element_text(size=8))

Rmisc::multiplot(g1, g4, g2,
                 g5, g3, g6,
                 cols = 3)

# z <- ggplot(data = who %>% mutate(past = ifelse(past, 'Historical',
#                                            'WHO objective')) %>%
#          filter(!is.na(with_malaria)) ,
#        aes(x = year,
#            y = with_malaria,
#            color = past)) +
#   geom_line() +
#   geom_point() +
#   theme_paper() +
#   labs(x = '',
#        y = 'Countries',
#        title = 'Countries with malaria: Observed and WHO objectives',
#        subtitle = 'Data from http://www.who.int/malaria/areas/elimination/overview/en/') +
#   scale_color_manual(name = '',
#                      values = c('darkorange', 'blue')) +
#   ylim(0, 130) +
#   theme(axis.text.x = element_text(angle = 90)) +
#   geom_line(data = projection,
#             aes(x = year,
#                 y = countries,
#                 color = NA,
#                 group = NA),
#             color = 'white',
#             lty = 2) +
#   geom_hline(yintercept = 0, col = 'red')
# 
# # xa_plot <- xa %>% mutate(year = year + 2016) %>% filter(year == 2066)
# xa_plot <- data.frame(year = c(2015, 2066),
#                       val = c(100, 0))
#   
# z + geom_line(data = xa_plot,
#               aes(x = year,
#                   y = val),
#               color = 'white',
#               alpha = 0.6,
#               group = 1) +
#   theme_black()
```

```{r}
x <- df %>%
  filter(responded) %>%
  group_by(n_areas) %>%
  tally %>%
  mutate(p = n / sum(n, na.rm = TRUE) * 100)
```

A total of `r length(which(df$responded))` researchers participated in the survey from the `r nrow(df)` invitations sent (participation rate of `r round(length(which(df$responded)) / nrow(df) * 100, 2)`%). Areas of expertise were non-exclusive and self-described, with participants having the option to choose from up to 3 of 10 checkboxes, or to write in one or more "other" areas of expertise.\footnote{Response categories were Anthropology, Biochemistry, Bioinformatics, Biology, Chemistry, Clinical medicine, Drug discovery, Ecology, Economics, Entomology, Epidemiology, Geography, GIS, History, Immunology, Infectious disease, IT, Malaria, Medical entomology, Medicinal chemistry, Microbiology, Parasitology, Pharmacology, Pharmacy, Political science, Public health, Statistics, Vector biology, Vector control, and Virology.} `r sum(x$n[1:3])` (`r round(sum(x$p[1:3]), digits = 1)`%) participants declared three or fewer areas of expertise.

```{r}
temp <- areas_df %>%
  arrange(desc(count))
temp$area <- trimws(temp$area, which = 'both')
names(temp) <- c('Area', 'Researchers')
temp$Area <- as.character(temp$Area)
temp$Area <- Hmisc::capitalize(temp$Area)
temp$Area[temp$Area == 'It'] <- 'Information Technology'
# kable(temp[1:10,]) #%>%
  # kableExtra::kable_styling(#position = "float_right",
  #                           font_size = 7)
```

Participants had a total of `r length(old_areas)` unique (self-described) areas of expertise. The five most popular were `r paste0(paste0(temp$Area[1:5], ' (', temp$Researchers[1:5], ')'), collapse = ', ')`.

\subsection*{Quantitative results}

Participants were asked to quantify (as a percentage) the likelihood of eradication in the next 10, 20, 30, 40, and 50 years. Most participants saw eradication as extremely unlikely in the next 10-30 years, but increasingly likely thereafter. Figure 2 shows the distribution of year-specific likelihood perceptions (panels A-E), as well as an illustration of how both likelihood and uncertainty grow over time (panel F).

Participants were asked to provide the number of years they believe it will take until eradication can be achieved. `r length(which(is.na(df$years[df$responded])))` (`r round(length(which(is.na(df$years[df$responded]))) / nrow(df) * 100, digits = 1)`%) were either blank or uninteligible, whereas `r length(which(!is.na(df$years)))` participants responded. Among respondants, `r length(which(df$years >= 50))` (`r round(length(which(df$years >= 50)) / length(which(!is.na(df$years))) * 100, digits = 1)` %) estimated that it would be 50 or more years until eradication.

```{r, fig.height = 3.5, fig.width = 4, eval = FALSE}
ggplot(data = df,
       aes(x = years)) +
  geom_histogram(fill = 'darkorange',
                 alpha = 0.6) +
  theme_paper(base_size = 10) +
  labs(x = 'Years',
       y = 'Responses',
       title = 'Perceived years until eradication',
       subtitle = '(Estimates of greater than 50 years reduced to 50)')
```

Respondents were qualitatively different from non-respondents. Importantly, the average number of total author-specific citations was `r round(mean(df$n_citations_total[df$responded]), digits = 2)` among respondents, but `r round(mean(df$n_citations_total[!df$responded]), digits = 2)` among non-respondents. This suggests a tendency for more senior or impactful researchers not to respond. When examining the number of average citations per article, the difference between respondents remained: `r round(mean(df$n_citations_average[df$responded]), digits = 2)` among respondents, and `r round(mean(df$n_citations_average[!df$responded]), digits = 2)` among non-respondents, highlighting the greater impact of non-respondents relative to respondents. Males responded at a greater rate (`r round(mean(df$responded[df$sex == 'male'], na.rm = T) * 100, digits = 2)`) than females (`r round(mean(df$responded[df$sex == 'female'], na.rm = T) * 100, digits = 2)`).

```{r}
# function to get response rate per categorical variable
response_rate <- function(variable = 'sex'){
  df$x <- df[,variable]
  out <- df %>%
    group_by(x) %>%
    summarise(responded = length(which(responded)),
              n = n()) %>%
    ungroup %>%
    mutate(p = responded / n * 100)
  return(out)
}
out_list <- list()
vars <- c('sex',
            # 'ethnicity',
            'bin_citations')
for (i in 1:length(vars)){
  this_var <- vars[i]
  x <- response_rate(variable = this_var)
  x$variable <- this_var
  out_list[[i]] <- x
}
out <- bind_rows(out_list) %>%
  dplyr::select(variable, 
                x,
                responded,
                n, 
                p) %>%
  rename(Variable = variable,
         Characteristic = x,
         Responded = responded,
         Invited = n,
         `Response rate` = p)
out$Variable <- Hmisc::capitalize(out$Variable)
out$Variable[out$Variable == 'Bin_citations'] <- 'Citations'
out$Characteristic <- Hmisc::capitalize(out$Characteristic)
out$`Response rate` <- paste0(round(out$`Response rate`, digits = 2), ' %')

for(i in nrow(out):2){
  this_var <- out$Variable[i]
  prev_var <- out$Variable[i-1]
  if(this_var == prev_var){
    out$Variable[i] <- ''
  }
}
knitr::kable(out, 
             # format = 'latex', 
             caption = 'Response rate by author characteristics')
```

Selection bias is not of concern in the case of differential response if the groups for whom there are differences are not different in terms of the outcome variable. This was the case for sex: males responded at a significantly greater rate than females (p < 0.001), but were not statistically significantly different in regards to pessimism/optimism (ie, time frame or likelihood of eradication). In the case of researcher impact (as measured by the total number of citations), selection bias plays an important role: being highly-cited was associated both with eradication "pessimism" as well as likelihood of non-response. In other words, our pool of respondents was less highly-cited than our pool of invitees, and among respondents, those who were highly-cited tended to be more pessimistic.

In order to de-bias sample selection, we employed Heckman's two-step correction method as well as simple binomial logistic regression model to estimate the likelihood of response as a function of sex and (binned) number of citations. To the extent that the two methods presented nearly identical value, we present here the results of the latter. Having estimated the odds of survey participation, we then used the inverse of the selection model's predictions as weights in a simple linear model to adjust our estimates. We run a separate weighted model to estimate the likelihood of eradication at 10, 20, 30, 40, and 50 years. Figure 3 shows both the aggregated perceived likelihood of eradication over time before and after adjusting for sample selection based on binned number of publications. Our adjusted and unadjusted estimates coincide in the long-term, but diverge sharply in the near-term.

```{r}
# Estimate model for de-biasing sample selection without Heckman.
response_model <- glm(responded ~ bin_citations + sex, data = df, family = binomial('logit'))
# Use model to predict likelihood of response
df$likelihood_of_response <- predict(response_model, df, type = 'response')
# Use inverse of likelihood of response to assign weight for model
df$weight <- 1 / df$likelihood_of_response
# Run likelihood of eradication models
for (i in seq(10, 50, 10)){
  assign(paste0('formula_simple_years_', i),
         as.formula(paste0('years_',
                           i,
                           '_numeric ~ sex')))

                           # '_numeric ~ sex + bin_citations')))
  # assign(paste0('simple_', i),
  #        lm(get(paste0('formula_simple_years_', i)), data = df))
  assign(paste0('simple_', i, '_weighted'),
         lm(get(paste0('formula_simple_years_', i)), data = df,
            weights = df$weight))

}

# Get aggregate adjusted vs unadjusted likelihoods of eradications
simple <- data_frame(years = seq(10, 50, 10),
                     adjusted = NA,
                     unadjusted = NA)
simple$unadjusted <- heckman$value[heckman$key == 'Unadjusted']
for (i in seq(10, 50, 10)){
   # simple$unadjusted[simple$years == i] <- mean(df[,paste0('years_', i, '_numeric')], na.rm = TRUE)
   simple$adjusted[simple$years == i] <- weighted.mean(x = df[,paste0('years_', i, '_numeric')],
                                                       w = ifelse(is.na(df$weight), 0,
                                                                  df$weight), na.rm = TRUE)
  # simple_model <- get(paste0('simple_', i))
  # weighted_model <- get(paste0('simple_', i, '_weighted'))

}
simple <- gather(simple, key, value, adjusted:unadjusted)
g2 <- ggplot(data = simple,
       aes(x = years,
           y = value,
           color = key,
           lty = key)) +
    geom_line(alpha = 0.8) +
  geom_point(alpha = 0.6) +
  theme_paper(base_size = 13) +
  labs(x = 'Years',
       y = 'Likelihood',
       subtitle = 'B') +
  scale_linetype_manual(name = '', values = 1:2) +
  scale_color_manual(name = '', values = c('blue', 'darkorange'))
x <- df %>%
  group_by(citations = bin_citations) %>%
  summarise(years_10 = mean(years_10_numeric, na.rm = TRUE),
            years_20 = mean(years_20_numeric, na.rm = TRUE),
            years_30 = mean(years_30_numeric, na.rm = TRUE),
            years_40 = mean(years_40_numeric, na.rm = TRUE),
            years_50 = mean(years_50_numeric, na.rm = TRUE),
            n = length(unique(email))) %>%
  ungroup %>%
  gather(key, value, years_10:years_50) %>%
  filter(!is.na(citations))

cols <- colorRampPalette(c('red', 'orange', 'blue') #brewer.pal(n = 9, 'Spectral')
                         )(length(unique(x$citations)))
g1 <- ggplot(data = x %>% filter(!is.na(citations)),
       aes(x = as.numeric(gsub('years_', '', key)),
           y = value,
           color = citations,
           group = citations)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.6) +
  theme_paper(base_size = 13) +
  scale_color_manual(name = 'N° citations',
                     values = cols) +
  labs(x = 'Years',
       y = 'Likelihood',
       subtitle = 'A')
```

```{r, fig.height = 3.5, fig.cap = 'Aggregated perceived likelihood of eradication over time, adjusted for sample selection and unadjusted.'}
# Rmisc::multiplot(g1, g2, cols = 2)
g2
```

Participants were heterogenous in their perceptions, with all of the 5 pre-determined cut-offs containing the whole range of likelihoods (0-100%), and the free-form time frame responses suggesting that eradication would be achieved in a time-frame ranging from less than a decade to never. Heterogeneity can only be partly explained by area of expertise, as the differences in perception between disciplines were minor and insignificant (see Figure 4).

```{r, fig.height = 3, fig.width = 4, fig.cap = 'Average perceived years to aggregation by discipline'}
areas_df <- areas_df %>% arrange(desc(count))
small_areas <- areas_df$area[1:12]
model_formula <- paste0('years ~ ',
                        paste0(paste0('area_', gsub(' ', '_', small_areas)), collapse = ' + '))
model_formula <- as.formula(model_formula)
fit <- lm(model_formula, data = df)
# Get confidence intervals too
ci <- tidy(confint(fit))
x <- cbind(tidy(fit), ci)
names(x) <- c('term', 'estimate', 'se', 'stat', 'p', 'rownames', 'lwr', 'upr')
x <- x %>%
  dplyr::select(term, estimate,
                p, lwr, upr)
the_intercept <- x$estimate[1]
x <- x[-1,]
for (the_column in c('estimate', 'lwr', 'upr')){
  x[,the_column] <- x[,the_column] + the_intercept
}
x$term <- gsub('area_|TRUE', '', x$term)
x <- x %>%
  dplyr::arrange(estimate)
x$term <- gsub('_', '\n', x$term)
x$term <- Hmisc::capitalize(x$term)
x$term[x$term == 'It'] <- 'Information\ntechnology'

x$term <- factor(x$term, levels = x$term)
ggplot(data = x,
       aes(x = term,
           y = estimate)) +
  geom_bar(stat = 'identity',
           fill = 'darkorange',
           alpha = 0.6) +
  geom_errorbar(aes(ymin = lwr,
                    ymax = upr),
                alpha = 0.6,
                width = 0.8,
                color = 'red') +
  theme_paper(base_size = 10) +
  labs(x = 'Area',
       y = 'Years') +
  theme(axis.text.x = element_text(angle = 90))

# # Alternative - median instead of mean
# alt <- data.frame(area = small_areas,
#                   med = NA,
#                   lwr = NA,
#                   upr = NA)
# for(i in 1:nrow(alt)){
#   message(i)
#   the_area <- alt$area[i]
#   the_area <- gsub(' ', '_', the_area)
#   the_column <- df[,paste0('area_', the_area)] * 100
#   alt$med[i] <- median(the_column, na.rm = TRUE)
#   alt$lwr[i] <- quantile(the_column, 0.25, na.rm = TRUE)
#   alt$upr[i] <- median(the_column, 0.75, na.rm = TRUE)
# }
# ggplot(data = alt,
#        aes(x = area,
#            y = med)) +
#   geom_bar(stat = 'identity', fill = 'darkorange') +
#   geom_linerange(aes(ymin = lwr,
#                      ymax = upr)) +
#   theme_paper()
```

```{r}
bb <- c('governance', 'health_workforce',
        'health_financing', 'health_technologies',
        'health_information', 'service_delivery')
results_list <- list()
for (i in 1:length(bb)){
  this_df <- data.frame(bb = bb[i],
                        ranking = as.numeric(df[,bb[i]]))
  results_list[[i]] <- this_df
}
results <- bind_rows(results_list)
results$bb <- gsub('_', '\n', results$bb)

results_agg <-
  results %>%
  dplyr::group_by(bb) %>%
  dplyr::summarise(avg = mean(ranking, na.rm = TRUE)) %>%
  arrange(avg)
results_agg$bb <- factor(results_agg$bb,
                         levels = results_agg$bb)
results$bb <- factor(results$bb, 
                     levels = levels(results_agg$bb))

# library(ggridges)
# ggplot(data = results %>% filter(!is.na(ranking)),
#        aes(x = ranking,
#            y = bb,
#            fill = bb)) +
#   geom_density_ridges(#fill = 'darkorange',
#                       alpha = 0.6,
#                       scale = 1.5) +
#   theme_paper() +
#   xlim(0,30) +
#   labs(x = 'Perceived importance',
#        y = 'WHO Building Block') +
#  scale_fill_cyclical(values = c("darkorange", "yellow"))

# ggplot(data = results) + 
#   geom_violin(aes(x = bb,
#            y = ranking),
#            fill = 'orange',
#               alpha = 0.6) +
#   geom_jitter(aes(x = bb,
#            y = ranking),
#            color = 'black',
#            with = 0.4,
#            height = 0.4,
#               alpha = 0.3,
#               size = 1) +
#   theme_paper(base_size = 10) +
#   labs(x = 'WHO Building block',
#        y = 'Ranking',
#        title = 'Perceived areas needing attention',
#        subtitle = 'Small points: jittered individual responses; large triangles: averages') +
#   geom_point(data = results_agg,
#              aes(x = bb,
#                  y = avg),
#              size = 5,
#              pch = 17,
#              color = 'black',
#              alpha = 0.8) +
#   ylim(0, 6)

# a <- data_frame(category = rep(c("Innovation",
#                                                    "Systemic Challenges",
#                                                    "Poverty",
#                                                    "Meta"),2),
#                 type = rep(c('Prescriptive', 'Descriptive'), each = 4),
#                 val = c(213,
#                         25,
#                         85,
#                         32,22,
#                         112,
#                         70,
#                         142))
# a$val <- ifelse(a$type == 'Prescriptive', a$val * -1, a$val)
# ggplot(data = a,
#        aes(x = category,
#            y = val)) +
#   geom_bar(alpha = 0.5,stat = 'identity', aes(fill = type)) +
#   coord_flip() +
#   geom_label(aes(label = abs(val))) +
#   theme(legend.position = 'bottom') +
#   scale_fill_manual(name = '',
#                     values = c('lightblue', 'darkorange')) +
#   theme_fivethirtyeight() +
#   ylim(-220, 150) +
#   labs(title = 'Classification of survey free responses',
#        subtitle = 'By thematic categorization and diagnostic')
```

## Qualitative analysis of comments

Of the `r length(which(df$responded))` who responded to the survey, `r length(which(!is.na(df$comments)))` (`r round(length(which(!is.na(df$comments))) / length(which(df$responded)) * 100, digits = 2)`%) provided a comment. Relative to non-commenters, commenters were more optimistic on average, but also more polarized in opinion. The three subject themes identified through iterative, open coding were:

1. *Solutions*: Comments pertaining to the innovations required to achieve eradication, priorities, and the desirability of certain approaches.

2. *Systemic challenges*: comments pertaining to political, social, environmental or logistical issues related to eradication.

3. *Complexity*: comments which focus on the multi-dimensional components of eradication.

Comments were also classified as descriptive or prescriptive. A majority (`r round(320 / 540 * 100, digits = 2)`%) were descriptive. Descriptive commenters were more pessimistic (in regards to perceived years until eradication) than prescriptive commenters, though this difference did not reach the level of statistical significance (p = 0.21). Descriptive comments also received sentiment polarity scores which were more negative than prescriptive comments, although again this difference did not reach the level of statistical significance (p = 0.18).

In regards to **solutions**, comments largely pertained to the necessity of further technological advances and innovations. One participant wrote that "currently available technology can't achieve [eradication], even if delivered optimally"; another argued that eradication could not be achieved without a "game-changing innovation", whereas multiple others referred to the need for "transformative" technologies:

\begin{quote}
We can't achieve eradication with our current tools. We'd need new medicines, a better vaccine, and maybe other vector control tools.
\end{quote}

Many noted the need to "overcome the challenge of drug resistance". More than 10% of commenters noted the need for an effective vaccine. Genetic engineering was mentioned by several commenters as a promising means to achieve eradication quickly. Most comments coded as "solutions" were presciptive in nature, often suggesting the nature of the needed innovation, with a heavy slant towards pharmaceutical options and vaccination.

**Systemic challenges** to malaria eradication were noted by the majority of commenters. Comments in this category can be divided into four sub-themes: (i) lack of coordination, (ii) lack of good surveillance and health services delivery, (iii) lack of political will and (iv) poverty. In direct contrast to the previous comments, many emphasized that "we already have the tools to achieve  eradication" and that the only piece lacking was "robust health systems". Many commenters noted problems of coordination, as illustrated in the below quote.

\begin{quote}
It will be very difficult to eradicate malaria... not because we don't have the technologies, which we already have... the problem is politics. Malaria doesn't stop in (sic) borders of a country and it would take a joint effort of a lot of political leaders to get programs in place to fight malaria. Unfortunately I don't see this happening anytime soon.
\end{quote}

Others echoed the sentiment, with many comments focusing on the need of strong surveillance and treatment delivery systems. Many commenters focused on other reasons for stagnating progress, such as "weak or failing health systems...due to political unwillingness or conflict".  Many noted that malaria is a "disease of poverty", with "social injustice" as the root cause. Some made the sequential argument that "eradication of poverty" must preceed disease eradication. Along the same lines, one commenter wrote:

\begin{quote}
Eradication requires a full systems-wide approach, not a disease-specific approach. The eradication of smallpox was a triumph of management, not medicine or technology.
\end{quote}

Another noted that the survey "left off the list the most important factor - economic development". Many echoed the sentiment, stating that "with no economic development, you cannot have eradication" and that poverty is the "cause" of malaria. Comments coded as "systemic" tended to be descriptive and more pessimistic than others.

**Complexity** was a relatively rare category (<20\% of all comments), those whose comments were coded as the "complexity" category were more pessimistic than average in regards to the timeline and likelihood of eradication. Many commenters highlighted the inherent challenges in the epidemiology of malaria, such as the changing dynamics of malaria transmission, the resilience of the parasite and vector, climate change, and the inability to aim interventions accurately with an "ever-moving target". The potential for adaptation was highlighted in reference both to the mosquito as well as the parasite itself. Many comments addressed the fact that the conversation on eradication is largely taking place within the public health community, whereas the causes of malaria endemicity are largely orthogonal to public health interventions, "going beyond the health sector". Several commenters pointed out the multitude of prerequisite conditions for eradication to even be considered feasible:

\begin{quote}
To my mind, this question is highly dependent on background context, e.g. global political and economic dynamics, as well as international conflict. Complete global eradication is an extremely singular goal that requires a vast array of necessary conditions - if any of these fail, eradication will not be achieved.  
\end{quote}

Many commenters thought that the terms "eradication" and "malaria" were so complex and nebulous that the public health practitioners should avoid the them all together, so as to not repeat the mistakes of the WHO GMP in the 1950s. For example, some highlighted that talking about "malaria" as one disease misses the mark, since the different species of parasite and contexts in which they live make elimination in each area very distinct from other areas.  One commenter wrote that eradication is a "postwar" idea that developed from the "abandonment of a broad sociopolitical understanding of the causes of disease, and the emphasis on technological solutions." Many stated that global malaria eradication was simply not possible, and two argued that it may not be desirable or ethical. One stated that the concept was "absurd" and that "I'm not even sure why people talk about it". Another questioned the utility of discussing "eradiction" as a concept:

\begin{quote}
Eradication is a different objective than elimination. Elimination means that the disease is not endemic but could reappear even in a country like Norway if infrastructure breaks down. Elimination may be possible in poor endemic countries, following socioeconomic development. Eradication means that the parasite disapears from the planet, which is not realistic...
\end{quote}

Comments questioning the utility of eradication as a concept or goal tended to be skeptical of its feasibility. Largely, they were prescriptive, advocating for a re-framing of the conversation so that the focus was not on an "arbitrary" goal like eradication, but rather on scaling up control and making region-specific progress.

# Discussion

Approximately three-quarters of respondents believe that malaria will not be eradicated in the next 50 years; in other words, given that our pool of respondents is of typical post-PhD age, most believe they will not live to see eradication. When adjusted for selection bias, the perceived likelihood of eradication 50 years from now remains similarly low, but estimates for shorter-term eradication are even lower.

In regards to differences by academic discipline, it is important to note that diversity within disciplines was greater than diversity between them. That said, one might surmise as to why disciplines at the extremes differed. Chemists, on average, reported that eradication will be achieved in less than 40 years, whereas economists thought it would take a full ten years more.  Speculatively, the collective optimism in the field of chemistry may be attributable to recent advances in drug development and insecticides, whereas economists may be more pessimistic due to concerns such as free-riding when it comes to global public goods [@barrett] [@Bodansky_2012]. This study was not designed to generate inference regarding reasons why disciplines might differ in perception, but it does suggest further lines of research.

Eradication of a disease is a "high stakes game", exposed to multiple competing - and at times contradictory - incentives from multiple stakeholders [@barrett]. Understanding these incentives, and the factors which can alter them, is vital to understanding how eradication can succeed, and where it is most vulnerable to failure. The return on investment of eradication-specific interventions is affected by the fact that most researchers agree that eradication will take a long time to achieve. This, in turn, reduces the expected value of future benefits, disincentivizing eradication-specific investment. Given this, it is important to quantify the positive externalities of "failed" eradication; in areas of high endemicity, even if the goal of elimination is not achieved, the reduction in burden or disease can still make interventions cost-effective.

However, areas of high malaria endemicity are often also those with high competing health costs. When estimating the return on investment of malaria elimination initiatives, not only must one take into account the potential benefits (even in the case of failure), but also the opportunity costs (even in the case of success). After all, elimination is not a binary success/failure proposition - an initiative should be judged both on its epidemiological circumstances [@Churcher_2014] as well as the counterfactual improvements in health which could have been achieved via other paths.

## Limitations

There are several reasons to be skeptical that the combined opinions of malaria researchers is a reliable indicator. First, academic researchers are specialists - their narrow, field-specific view of eradication's feasibility is of arguable reliability, given that they may be unfamiliar with the operational, cultural and "real-world" challenges of malaria eradication. Though crowds have been found to be more "wise" than individuals in many cases, the application of an approach similar to the "wisdom of crowds" is not suitable to all classes of problems [@Mannes_2014]. Crowds can be susceptible to social biases [@Lorenz_2011] (although this survey's anonimity largely protects against this issue), and other biases may come in to play, especially given that our study was of a crowd of "specialists", rather than the population as a whole.

Four potential biases are worth mentioning specifically. (1) "Conjunction fallacy" suggests that the general goal of eradication may seem less likely than the sum of the goals of country-specific elimination. (2) A (reverse) variant of the "hot hand fallacy", in which researchers may mistakenly base their assessment of current chances of eradication on previous failures. (3) Parkinson's law of triviality suggests that researchers may disproportionately see the challenges of their own research (antimalarial drug resistance, etc.) as larger or more relevant to the global eradication campaign than they really are. (4) Finally, and ironically, "optimism bias" may play a perverse role in researchers' responses; though eradication is certainly a goal desired by all, one could argue that malaria research specialists subconciously realize that they actually stand to lose out profesionally in the case of eradication.

As with Keenan et al's survey of experts regarding the feasibility of eradication of neglected tropical diseases [@Keenan2013], our survey did not delve into whether researchers had clinical or operational experience, nor did it assess opinions of program workers, nor did it explore complexities pertaining to different types of or forms of existence of malaria.

This study included the first authors of indexed journals. Though certainly a group with important knowledge related to malaria, this misses malaria control program employees, health agency workers, and other stakeholders. Their experiences and viewpoints may be different from those of academics, and arguably more relevant. Additionally, in public health journals, first authors are often more junior than their collaborators. By focusing solely on first authors, we may have unintentionally created a sample which is qualitatively different than the true universe of malaria researchers; to the extent that in our result pool researchers with less experience (as represented through publications) tended to be more "optimistic" regarding eradication, it is reasonable to assume that the restriction of first authors may have lead to an overly optimistic sample, making the results of the survey even more striking.

## Conclusion

Our survey attempted to quantify the likelihood of, and time frame to, malaria eradication by gauging the collective opinion of malaria researchers.  The findings of our survey show researchers expressing hesitance about the likelihood of eradication and suggesting a long time frame. The causes for scepticism are diverse, but common themes were the need for innovation, systemic challenges, and the complexity of the disease and its transmission.

The implication of these results are two-fold: (1) that those working or investing in eradication-specific campaigns, as well as those modelling these campaigns' hypothetical cost-effectiveness, should factor in researchers' perceived low likelihood and long time frame when calculating those campaigns' expected value; (2) that champions of near-term eradication may need to make a more compelling case to malaria researchers of eradication's feasibility, in order to better focus and inspire the latter.

We have strived to present the results of our study without insinuation regarding the "true" feasibility and timeframe of eradication, as only time will tell whether the collective "wisdom" of researchers was worth adhering to or not. The actual cost-effectiveness of eradication interventions is not only a function of eradication's success, but also of a number of other factors which are only knowable retrospectively. This study's primary contribution is the provision of a snapshot of perceptions of malaria researchers, whose opinions may be of value not only to other researchers, but also to the malaria and public health communities at large.

References {#references .unnumbered}
==========